The Power of Two Random Choices

### 2014-03-31 presentation/presentation.tex
rebaseme
use symlinks for source files
some more polishing after tips from francesco

\documentclass[serif,professionalfonts]{beamer}
\usetheme{Forcellini}

\usepackage[backend=bibtex]{biblatex} 
\addbibresource{../sources.bib}

@@

% Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
\renewcommand{\sfdefault}{uop}
\renewcommand{\rmdefault}{ppl}

@@

\setbeamertemplate{caption}
{
  \raggedright%
  \insertcaption\par%
}
\setbeamercolor{green}{fg=green}

\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
\newcommand\load{L_{\mathrm{max}}}
\newcommand\loadS{\load^{\mathrm{S}}}
\newcommand\loadG{\load^{\mathrm{G}}}
\newcommand\loadAgl{\load^{\mathrm{A}}}
\newtheorem{claim}{Claim}
\newtheorem{assumption}{Assumption}
\newtheorem{invariant}{Invariant}

@@

%---------------------------------------------------------------------
%Front matter
%---------------------------------------------------------------------
\title{The Power of Two Random Choices}
\subtitle{\dots and the Surprising Influence of Asymmetry}
\author[M. Kalany]{Martin Kalany}
\institute[TU Wien]
{
  Graduate student in Computer Science\\
  Vienna University of Technology\\
}
\date{\today}

@@

\begin{frame}
\frametitle{The problem}
\framesubtitle{...and the goal}
\begin{exampleblock}{Balls-into-bins games}
Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls among them, \alert{as evenly as possible}. 
\end{exampleblock}
\pause
\begin{exampleblock}{Definition}
We are interested in the \alert{maximum load $\load$}, i.e, how many balls are in the fullest bin?
\end{exampleblock}

@@

\medskip
\begin{theorem}[Raab and Steger, 1998]
When using the \alert{single choice} strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadS \leq 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\bigskip
We distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
\begin{enumerate}
\item Uniformly and independently
\item Non-uniformly and independently
\item Non-uniformly and dependently
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{The Greedy algorithm}
\begin{exampleblock}{Algorithm}
\begin{enumerate}
\item For each ball, choose $d$ bins independently and uniformly at random. 
\item Put the ball into the least full bin.
\item In case a tie occurs, choose one of the $d$ bins arbitrarily.
\end{enumerate}
\end{exampleblock}

@@

\only<2->{
\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
\end{align*}
\end{theorem}
}
\end{frame}

\begin{frame}
\frametitle{Single choice vs.~Greedy}
\begin{align*}
\loadS \leq 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases}
\end{align*}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}

\begin{alertblock}{The two-choice paradigm}
Even for $d=2$, Greedy achieves an \alert{exponential decrease} of the additive gap $\delta$, which does \alert{not depend on} the number of balls \alert{$m$}.
\end{alertblock}
\end{frame}

\begin{frame}
\frametitle{Applications}
\begin{description}
	\item[Hashing:] Assuming a perfect hash function, $n$ elements (balls) are mapped to $m$ table entries (bins). Using two hash functions: look-up time $\Theta(\ln\ln n)$, w.h.p. \alert{Easy to parallelize} and  \alert{no re-hashing}!
	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins). The two-choice paradigm helps to \alert{reduce communication overhead}, %(of e.g., a central dispatcher)
	while the maximum load will be $\Theta(\ln\ln n)$, w.h.p.
	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM s.t.~the overhead of the emulation is minimized.
	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} %(a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination) 
	is analogous to the single-choice technique.
\end{description}
\end{frame}

\newcommand\aglAlgorithm{
\begin{exampleblock}{Algorithm}
\begin{enumerate}
\item Partition the $n$ bins into $d$ groups.
\item Assume a fixed ordering of the bins $b_i$, $1\leq i \leq n$.
\item Choose one bin from each group uniformly and independently.
\item Put the ball into the least full bin.
\item In case a tie occurs, put the ball into the leftmost bin (i.e., the bin with smallest index) among those with same minimal load.
\end{enumerate}
\end{exampleblock}
}

@@

\begin{frame}
\frametitle{$d$-step Fibonacci numbers and the golden ratio}
\framesubtitle{(Capocelli, Cull, 1990)}

$d$-step Fibonacci numbers:
\begin{align*}
F_d(k) := \begin{cases}
                0               & k \leq 0\\
                1               & k = 1\\
                \sum_{i=1}^{d}F_d(k-i) & k > 1
            \end{cases}
\end{align*}
Generalized golden ratio $\Phi_d$:
\begin{align*}
\Phi_d := \lim_{k \rightarrow \infty} \frac{F_d(k)}{F_d(k-1)}
\end{align*}
\begin{align*}
\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 
\end{align*}
\begin{align*}
F_d(k) \geq \Phi_d^{k-2}
\end{align*}
\end{frame}

\newcommand\theoremVocking{
\begin{theorem}[V\"ocking, 2003]
When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}
}

\begin{frame}[shrink]
\frametitle{The Always-go-left algorithm}
\aglAlgorithm
\theoremVocking
\end{frame}

\begin{frame}
\frametitle{The Always-go-left algorithm}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}
\end{column}
\begin{column}{0.5\textwidth}
\begin{align*}
\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
\end{column}
\end{columns}
\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
\bigskip
\begin{theorem}[V\"ocking, 2003]
For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3), the maximum load is, w.h.p,
\begin{align*}
\loadAgl = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
\end{align*}
\end{theorem}
\end{frame}

\begin{frame}
\frametitle{BiB Model}
We will assume a sequential, infinite, dynamic process.

\begin{description}
\item[Sequential:] The balls arrive one after another, one at each time step~$t$.
\item[Infinite:] Ball placements and deletions are given for an infinite time span.
\item[Dynamic:] Balls do not stay in the bins forever.
\begin{itemize}
\item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
\item $d_t$/$a_t$: ball to delete/add at time step t. \\
\item $\sigma$ is fixed in advance.
\end{itemize}
\end{description}
\end{frame}

\begin{frame}
\frametitle{The Always-go-left algorithm}
\framesubtitle{Proof of the upper bound for $m=n$}
\begin{theorem}[V\"ocking, 2003]
\label{theorem:aglm}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Always-go-left algorithm, the maximum load at any time step $t$ is, w.h.p., 
\begin{align*}
\loadAgl \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}

\pause
\bigskip
In case $\load$ exceeds a given threshold of $L+3$ balls ("a bad event occurs"), we construct an activated \alert{witness tree of order $L$}. The probability for its existence upper bounds the probability that the associated bad event occurs.  
\end{frame}

@@

\newcommand\defEdgeEvent{
\begin{exampleblock}{Edge event}
Edge $e =(u,v)$, s.t.~$v$ is $i$-th child of $u$: $e$ represents the event that the $i$-th location of ball $b_u$ points to the same bin as one of the locations of ball $b_v$.
\end{exampleblock}
}

\newcommand\defLeafEvent{
\begin{exampleblock}{Leaf event}
A leaf node $v$ represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls, at the time of the insertion of the ball $b_v$.
\end{exampleblock}
}

\begin{frame}
\frametitle{Edge events}
\defEdgeEvent
\begin{center}
\begin{tikzpicture}[every tree node/.style={thick}]
 \Tree 
 	[.\node[draw,circle]{u}; 
		\edge[draw, densely dotted]; {}  
		%\edge[draw, densely dotted]; {}
		\edge[draw] node[draw=none, thick, anchor=west] {$i$};
		[.\node[draw,circle]{v};
		]
	]
	\node (bu) at (0.75,0.5) {};
	\node (bv) at (1.1,-1.25) {};

@@

\begin{frame}[shrink]
\frametitle{Asymmetric witness trees (AWT)} 
We assign a label $(h,i)$, with $0\leq h \leq L$, $1\leq i \leq d$, to each node:
\begin{itemize}
\item Node $T_d(1)$/$T_d(2)$ is labeled $(0,1)$/$(0,2)$.
\item For $i>2$, a node $(0,i)$ has $i-1$ children $(0,i-1),\dots,(0,1)$.
\item For $h>0$, a node $(h,i)$, has $d$ children
$(h,i-1),\dots,(h,1)$, $(h-1,d),\dots,(h-1, i)$.
\item Thus, the node $T_d(dh+i)$ is labeled $(h,i)$.
\end{itemize}
\begin{center}
\begin{tikzpicture}
 \Tree 
	[.$T_d(6)$\\$(1,3)$ \edge[draw, dotted];
		[.$T_d(5)$\\$(1,2)$
			[.$T_d(4)$\\$(1,1)$
				[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
				$T_d(2)$\\$(0,2)$ %$T_d(1)$\\$(0,1)$
			]
			[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
			%$T_d(2)$\\$(0,2)$
		]
	\edge[draw=none]; {}
	]
\end{tikzpicture}
\end{center}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
\begin{block}{Lemma}
The occurrence of a bad event (i.e., a bin contains more than $L+3$ balls) implies the activation of an AWT of order $L$.
\end{block}
\bigskip
\visible<2>{
We assign balls to the nodes of the AWT s.t.:
\myInvariant
}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
Given a node $v$ with label $(h,i)$ that references a ball $b_v$ (which was put into some bin of group $i$ that contained at least $h+3$ other balls at the time of insertion of ball $b_v$), \alert{we conclude} that:
\begin{enumerate}
\only<1>{
\item For $1\leq j < i$, the $j$-th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$. If it contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$. We assign the topmost ball in location $j$ to the child $u$ of $v$ with label $(h,j)$. 
}
\only<2>{
\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls. If it contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to. We assign the topmost ball in location $j$ to the child node $u$ of $v$ labeled with $(h-1, j)$. \qed

}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Probability of activation}
\begin{lemma}
For any $\alpha > 0$ and large enough $L$, the activation of an AWT of order L occurs with probability at most $n^{-\alpha}$.
\end{lemma}
\medskip
\begin{itemize}
\item $k$\dots number of nodes
\item $l$\dots number of leaves
\item $\omega$\dots number of ways to assign balls to the nodes of an AWT
\item $p_\mathrm{e}$\dots probability that all edge events occur
\item $p_\mathrm{l}$\dots probability that all edge events occur
\item $P_\mathrm{a}$\dots probability of activation of an AWT of order $L$
\end{itemize}

>>>>

\documentclass[professionalfonts]{beamer}
\usepackage{pgfpages}

\usetheme{Forcellini}

\usepackage[backend=bibtex]{biblatex} 
\addbibresource{sources.bib}

@@

% Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
%\renewcommand{\sfdefault}{uop}
%\renewcommand{\rmdefault}{ppl}
\usepackage{eulervm}

@@

\setbeamercolor{green}{fg=green}

\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}

\DeclareMathOperator{\load}{L_{\max}} % ensuremath
\DeclareMathOperator{\loadS}{L_{\max}^{S}} 
\DeclareMathOperator{\loadG}{L_{\max}^{G}} 
\DeclareMathOperator{\loadAgl}{L_{\max}^{Agl}} 
\DeclareMathOperator{\deltaS}{S}
\DeclareMathOperator{\deltaG}{G}
\DeclareMathOperator{\deltaAgl}{Agl}

@@

%---------------------------------------------------------------------
%Front matter
%---------------------------------------------------------------------
\title{The Power of Two Random Choices}
\subtitle{\dots and the surprising influence of asymmetry}
\author[M. Kalany]{Martin Kalany}
\institute[TU Wien]
{
  Graduate student in Computer Science\\
  Vienna University of Technology\\
}
\date{\today}

@@

\begin{frame}
\frametitle{The problem}
\framesubtitle{...and the goal}
\begin{alertblock}{Balls-into-bins games}
Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls
among them, \alert{as evenly as possible}. 
\end{alertblock}
\pause
\begin{exampleblock}{Definition}
We are interested in the \alert{maximum load $\load(m,n)$}, i.e, how many balls are in the fullest bin?
\end{exampleblock}

@@

\medskip
\begin{theorem}[Raab and Steger, 1998]
When using the \alert{single choice} strategy to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadS = 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n}  \, + \, \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\bigskip
We distinguish placement algorithms by the way the \alert{$d$ bins are sampled}:
\begin{enumerate}
\item Uniformly and independently
\item Non-uniformly and independently
\item Non-uniformly and dependently
\end{enumerate}
\end{frame}

\begin{frame}[t]
\frametitle{The Greedy algorithm}
\begin{exampleblock}{Algorithm}
\begin{enumerate}
\item For each ball, choose $d$ bins independently and uniformly at random. 
\item Put the ball into the least full bin.
\item In case a tie occurs, choose one of the $d$ bins arbitrarily.
\end{enumerate}
\end{exampleblock}

@@

\only<2->{
\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
When using the \alert{Greedy} algorithm to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
\end{align*}
\end{theorem}
}
\end{frame}

\begin{frame}
\frametitle{Single choice vs.~Greedy}
\begin{align*}
\delta^{\deltaS} &= 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases}\\[5mm]
\delta^{\deltaG} &\leq\frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}

\begin{alertblock}{The two-choice paradigm}
  Even for $d=2$, Greedy's additive gap $\delta$
  \begin{itemize}
  \item decreases \alert{exponentially} compared to the single choice algorithm,
  \item and does \alert{not depend on} the number of balls \alert{$m$}.
  \end{itemize}
\end{alertblock}
\end{frame}

\begin{frame}
\frametitle{Applications}
\begin{description}
	\item[Hashing:] Assuming a perfect hash function, $m$ elements (balls) are mapped to $n$ table entries (bins). Using two hash functions: look-up time $O(\ln\ln n)$, w.h.p. \alert{Easy to parallelize} and  \alert{no re-hashing}!
	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins). The two-choice paradigm helps to \alert{reduce communication overhead}, %(of e.g., a central dispatcher)
	while the maximum load is $O(\ln\ln n)$, w.h.p.
	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM, s.t.~the overhead of the emulation is minimized.
	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} (first
          route to some random intermediate
          destination) %(a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination)
          is analogous to the single-choice technique.
\end{description}
\end{frame}

\newcommand\aglAlgorithm{
\begin{exampleblock}{Algorithm}
\begin{enumerate}
\item \alert{Partition} the $n$ bins into $d$ groups.
\item Assume a \alert{fixed ordering} of the bins $b_i$, $1\leq i \leq n$.
\item Choose one bin \alert{from each group} uniformly and independently.
\item Put the ball into the \alert{least full} bin.
\item In case a \alert{tie} occurs, put the ball into the \alert{leftmost} bin (i.e., the bin with smallest index) among those with same minimal load.
\end{enumerate}
\end{exampleblock}
}

@@

\begin{frame}
\frametitle{The $d$-step Fibonacci numbers and the golden ratio}
\framesubtitle{(Capocelli and Cull, 1990)}

\begin{itemize}
\item The \alert{$d$-step} Fibonacci numbers:
  \begin{align*}
    F_d(k) := \begin{cases}
      0               & k \leq 0\\
      1               & k = 1\\
      \sum_{i=1}^{d}F_d(k-i) & k > 1
    \end{cases}
  \end{align*}
\item Generalized \alert{golden ratio} $\Phi_d$:
  \begin{align*}
    \Phi_d := \lim_{k \rightarrow \infty} \frac{F_d(k)}{F_d(k-1)}
  \end{align*}
  \begin{align*}
    \Phi_2 \approx 1.618 \  , \ \forall i\; \forall j \in \mathbb{N}: i < j
    \implies \Phi_i < \Phi_j \ , \ \lim_{d\rightarrow \infty} \Phi_d = 2
  \end{align*}
  \begin{align*}
    \forall d\; \forall k\geq	2: \quad F_d(k) \geq \Phi_d^{k-2}
  \end{align*}
\end{itemize}
\end{frame}

\newcommand\theoremVocking{
\begin{theorem}[V\"ocking, 2003]
When using the \alert{Always-go-left} algorithm to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}
}

\begin{frame}
\frametitle{The Always-go-left algorithm}
\aglAlgorithm
\theoremVocking
\end{frame}

\begin{frame}
\frametitle{The Always-go-left algorithm}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{align*}
\delta^{\deltaG} \leq \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}
\end{column}
\begin{column}{0.5\textwidth}
\begin{align*}
\delta^{\deltaAgl} \leq \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
\end{column}
\end{columns}
\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
\bigskip
\begin{theorem}[V\"ocking, 2003]
For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random, the maximum load will be, w.h.p,
\begin{align*}
\loadAgl = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
\end{align*}
\end{theorem}
\end{frame}

\begin{frame}{Balls-into-bins model: formal definition}
We will assume a sequential, infinite, dynamic process.

\begin{description}
\item[Sequential:] The balls arrive one after another, one at each time step~$t$.
\item[Infinite:] Ball placements and deletions are given for an infinite time span.
\item[Dynamic:] Balls do not stay in the bins forever.
\begin{itemize}
\item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
\item $d_t$ (and $a_t$): ball to delete (and add) at time step t. \\
\item $\sigma$ is fixed in advance.
\end{itemize}
\end{description}
\end{frame}
\end{comment}

\begin{frame}
\frametitle{The Always-go-left algorithm}
\framesubtitle{Proof of the upper bound for $m=n$}
\begin{theorem}[V\"ocking, 2003]
\label{theorem:aglm}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into the bins by the Always-go-left algorithm, the maximum load at any time step $t$ is, w.h.p., 
\begin{align*}
\loadAgl \leq \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}

\bigskip
In case $\load$ exceeds a given threshold of $L+3$ balls ("a bad event occurs"), we construct an activated \alert{witness tree of order $L$}. The probability for its existence upper bounds the probability that the associated bad event occurs.  
\end{frame}

@@

\newcommand\defEdgeEvent{
\begin{exampleblock}{Edge event}
Edge $e =(u,v)$, s.t.~$v$ is $i$-th child of $u$: $e$ represents the event that the $i$-th location of ball $b_u$ points to the \alert{same bin} as one of the locations of ball $b_v$.
\end{exampleblock}
}

\newcommand\defLeafEvent{
\begin{exampleblock}{Leaf event}
A leaf node $v$ represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains \alert{at least three additional} balls, at the time of the insertion of the ball $b_v$.
\end{exampleblock}
}

\begin{frame}
\frametitle{Edge events}
\defEdgeEvent
\begin{center}
\begin{tikzpicture}[every tree node/.style={thick}]
 \Tree 
 	[.\node[draw,circle]{u}; 
		\edge[draw, densely dotted]; {}  
		\edge[draw] node[draw=none, thick, anchor=west] {$i$};
		[.\node[draw,circle]{v};
		]
	]
	\node (bu) at (0.75,0.5) {};
	\node (bv) at (1.1,-1.25) {};

@@

\begin{frame}
\frametitle{Asymmetric witness trees (AWT)} 
We assign a label $(h,i)$, with $0\leq h \leq L$, $1\leq i \leq d$, to each node:
\begin{itemize}
\item Node $T_d(1)$/$T_d(2)$ is labeled $(0,1)$/$(0,2)$.
\item For $i>2$, a node $(0,i)$ has $i-1$ children $(0,i-1),\dots,(0,1)$.
\item For $h>0$, a node $(h,i)$, has $d$ children
$(h,i-1),\dots,(h,1)$, $(h-1,d),\dots,(h-1, i)$.
\item Thus, the node $T_d(dh+i)$ is labeled $(h,i)$.
\end{itemize}
\begin{center}
\begin{tikzpicture}
 \Tree 
	[.$T_d(6)$\\$(1,3)$ \edge[draw, dotted];
		[.$T_d(5)$\\$(1,2)$
			[.$T_d(4)$\\$(1,1)$
				[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
				$T_d(2)$\\$(0,2)$ %$T_d(1)$\\$(0,1)$
			]
			[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
			%$T_d(2)$\\$(0,2)$
		]
	\edge[draw=none]; {}
	]
\end{tikzpicture}
\end{center}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
\begin{block}{Lemma}
The occurrence of a \alert{bad event} (i.e., a bin contains more than $L+3$ balls) implies the \alert{activation} of an AWT of order $L$.
\end{block}
\bigskip
\visible<2>{
We assign balls to the nodes of the AWT s.t.:
\myInvariant
}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
Given a node $v$ with label $(h,i)$ that references a ball $b_v$ (which was put into some bin of group $i$ that contained at least $h+3$ other balls at the time of insertion of ball $b_v$), \alert{we conclude} that:
\begin{itemize}
\item For $1\leq j < i$, the $j$-th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$. If it contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$. We assign the topmost ball in location $j$ to the child $u$ of $v$ with label $(h,j)$. 
\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls (by similar arguments as above). We assign the topmost ball in location $j$ to the child node $u$ of $v$ labeled with $(h-1, j)$. \qed
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Probability of activation}
\begin{lemma}
For any $\alpha > 0$ and large enough $L$, the activation of an AWT of order L occurs with probability at most $n^{-\alpha}$.
\end{lemma}
\medskip
\begin{itemize}
\item $k$: number of nodes
\item $l$: number of leaves
\item $\omega$: number of ways to assign balls to the nodes of an AWT
\item $p_\mathrm{e}$: probability that all edge events occur
\item $p_\mathrm{l}$: probability that all edge events occur
\item $P_\mathrm{a}$: probability of activation of an AWT of order $L$
\end{itemize}
